---
title: "Run nowcasting approaches on real-world line list data"
---
# Preparation
## Setup
```{r, message=FALSE, include=FALSE}
here::i_am("code/nowcast_switzerland_hosp_symp.Rmd")
source(here::here("code", "utils", "setup.R"))
source(here::here("code", "utils", "utils_job.R"))
source(here::here("code", "utils", "utils_cluster.R"))
source(here::here("code", "utils", "utils_simulate.R"))
```

## Data loading
Line list data are available under terms of data protection upon request from the Swiss Federal Office of Public Health (FOPH).
```{r}
# get the preprocessed data
prep_data <- readRDS(here::here("data", "preprocessed", paste0("prep_symp-rep_", "DPTv2", ".rds")))
latest_date <- prep_data %>% pull(event2_date) %>% max()
earliest_date <- prep_data %>% pull(event2_date) %>% min()
print(paste("Using data until", latest_date))
```

## Exploratory analysis

### Case counts
Comparison of cases with known and missing symptom onset date
```{r, fig.width = 10. fig.height = 6}
(prep_data %>%
  mutate(so = factor(ifelse(!is.na(event1_date), "Known", "Missing"),
                     levels = c("Missing", "Known"), ordered = T)) %>%
  ggplot(aes(x = event2_date, fill = so)) +
  geom_bar(position = "stack") +
  theme_bw() +
  xlab("Date of report") +
  ylab("Case count") +
  scale_fill_manual(values = c("#d9d9d9", "#999999"), name = "Symptom onset") +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_date(date_breaks = "4 month", expand = c(0, 0)) +
  coord_cartesian(xlim = as.Date(c("2020-03-01", "2021-03-31"))) +
  theme(legend.position = "top"))

ggsave(here::here("figures", "paper", "switzerland_date_of_report.pdf"), height = 5, width = 10)
```

### Reporting delay

Empirical delay and maximum delay
```{r}
(prep_data %>%
  mutate(delay = as.integer(event2_date - event1_date)) %>%
  count(delay) %>%
  filter(!is.na(delay), delay >= 0) %>%
  mutate(p = n / sum(n)) %>%
  ggplot(aes(x = delay, y = p)) +
  geom_col() +
  geom_vline(xintercept = c(7 * 8, 7 * 12), color = "red") +
  scale_y_continuous(labels = scales::percent) +
  coord_cartesian(xlim = c(0, 200)) +
  theme_bw() +
  ggtitle("Reporting delay distribution"))

(prep_data %>%
  mutate(delay = as.integer(event2_date - event1_date)) %>%
  count(delay) %>%
  filter(!is.na(delay), delay >= 0) %>%
  mutate(p = cumsum(n / sum(n))) %>%
  ggplot(aes(x = delay, y = p)) +
  geom_col() +
  geom_vline(xintercept = c(7 * 8, 7 * 12), color = "red") +
  scale_y_continuous(labels = scales::percent) +
  coord_cartesian(xlim = c(0, 200)) +
  theme_bw() +
  ggtitle("Cumulative reporting delay distribution"))

print(paste0("A reporting delay of 8 weeks covers ", prep_data %>% mutate(delay = event2_date - event1_date) %>% summarize(100 * sum(delay <= 7 * 8, na.rm = T) / sum(!is.na(delay))) %>% pull() %>% round(2), "% of cases."))
```

Empirical reporting delay in different periods
```{r}
pre_data_periods <- prep_data %>%
  mutate(delay = event2_date - event1_date) %>%
  filter(
    event1_date >= "2020-03-01",
    event1_date <= "2021-03-31",
    wday(event1_date, label = T) == "Mon"
  ) %>%
  mutate(period = ifelse(event1_date < "2020-06-01",
    "March 1, 2020 - May 31, 2020",
    ifelse(event1_date < "2020-10-01",
      "June 1, 2020 - Sep 31, 2020",
      "Oct 1, 2020 - March 31, 2021"
    )
  )) %>%
  mutate(period = factor(period,
    levels = c(
      "March 1, 2020 - May 31, 2020",
      "June 1, 2020 - Sep 31, 2020",
      "Oct 1, 2020 - March 31, 2021"
    ), ordered = T
  )) %>%
  group_by(period)

(pre_data_periods %>%
  count(delay) %>%
  filter(!is.na(delay), delay >= 0) %>%
  mutate(p = n / sum(n)) %>%
  ggplot(aes(x = delay, y = p)) +
  geom_col(fill = "darkgrey") +
  geom_vline(xintercept = c(7 * 8, 7 * 12), color = "red") +
  theme_bw() +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(hjust = 0)
  ) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = expansion(add = c(0, 0.01))) +
  coord_cartesian(xlim = c(0, 55)) +
  facet_wrap(~period) +
  ggsci::scale_fill_npg() +
  xlab("Empirical forward reporting delay [days] on Mondays") +
  ylab("PMF"))

ggsave(here::here("figures", "paper", "switzerland_empirical_delay.pdf"), height = 3, width = 10)
```

Corresponding delay statistics in different periods
```{r}
pre_data_periods %>%
  filter(delay < 7 * 8) %>%
  summarize(across(delay, list(mean = mean,
                               median = median,
                               sd = sd), na.rm = T)) %>%
  mutate(across(starts_with("delay"), round, 2))
```

Mean empirical delay over time
```{r}
(prep_data %>%
  filter(!is.na(event1_date)) %>% 
  mutate(delay = as.integer(event2_date - event1_date)) %>% 
  group_by(event2_date) %>% summarize(mean_delay = mean(delay)) %>% 
  ggplot(aes(x = event2_date, y=mean_delay)) +
  geom_point() +
  geom_smooth(method = "loess", formula = y ~ x, span = 2) +
  theme_bw())
```

### Missing symptom onset date statistics
```{r}
prep_data_raw <- readRDS(here::here("data", "preprocessed", paste0("prep_symp-rep_", "DPTv2_raw", ".rds")))

# number of cases with negative delay (will be set to missing for nowcasting)
prep_data_raw %>% 
  mutate(delay = as.integer(event2_date - event1_date)) %>% 
  filter(!is.na(delay)) %>% 
  count(delay<0) %>%
  mutate(p = n/sum(n))

# overall number of cases
prep_data_raw %>%
  filter(event2_date <= "2021-03-31") %>%
  count()

# count missing / not missing
prep_data_raw %>%
  filter(event2_date <= "2021-03-31") %>%
  count(is.na(event1_date)) %>%
  mutate(p = n/sum(n))

# Plot with alpha over time
(prep_data_raw %>%
    filter(event2_date <= "2021-03-31") %>% 
    count(event2_date,missing=is.na(event1_date)) %>% 
    group_by(event2_date) %>%
    summarize(alpha = 1-sum(missing*n)/sum(n)) %>% 
  ggplot(aes(x = event2_date, y=alpha)) +
  geom_point() +
    geom_smooth(method = "loess", formula = y ~ x) +
    tidyquant::geom_ma(n = 14, size = 1, linetype = "solid", color = "red") +
    theme_bw() +
    xlab("Date of report") +
    ylab("Share of known onset dates") +
    scale_y_continuous(n.breaks = 20, labels = scales::percent, limits = c(0, NA), expand = c(0,0))
  )

# Missing by moving average over week
df <- prep_data_raw %>%
  filter(event2_date <= "2021-03-31") %>% 
  group_by(event2_date) %>% 
  summarize(nmissing=sum(is.na(event1_date)), nall=n(), .groups = "drop") %>% 
  arrange(event2_date)

summary(as.vector(
  zoo(df[,c("nmissing","nall")], order.by = df$event2_date) %>% 
  zoo::rollapply(width = 7,
                 FUN = function(x) sum(x[,"nmissing"])/sum(x[,"nall"]),
                 by.column = F,
                 align = "left")
  ))
```

# Nowcasting

Define incubation period and generation interval distribution over time
```{r}
source(here::here("code", "config_epi_params.R"))
plot_dist_ts_quantiles(latent_delay_dist_all) + ggtitle("Incubation period")
plot_dist_ts_quantiles(generation_time_dist_all) + ggtitle("Generation time")
```

Define nowcast dates / windows
```{r}
now <- seq.Date(as.Date("2021-03-31"), as.Date("2020-03-01"), by = "-1 day")
start_date <- now - 7 * 4 * 3 # cover last 3 months 
if(length(now)==1) print(paste0("Covering the period from ", start_date, " to ", now, "."))

# select relevant delay dists
latent_delay_dist_list <- latent_delay_dist_all_list[as.Date(names(latent_delay_dist_all_list)) %in% now]
generation_time_dist_list <- generation_time_dist_all_list[as.Date(names(generation_time_dist_all_list)) %in% now]

stopifnot(length(now)==length(latent_delay_dist_list))
```

## Job definition
Create nowcast job(array)
-- shared settings
```{r}
source(here::here("code", "utils", "setup.R"))

data_def <- define_data(list(prep_data), now, start_date, holidays = holidays_CHE)

sampling_def <- define_sampling(
  iter_warmup = 1000,
  iter_sampling = 1000,
  chains = 4,
  parallel_chains = 4,
  adapt_delta = 0.999,
  step_size = 0.01,
  max_treedepth = 15,
  inits_function = default_inits,
  refresh = 200,
  show_messages = T,
  temp_save = F,
  seed = 2)

output_def <- define_output(
  stan_data_list = FALSE,
  stan_prep_info = FALSE,
  priors = FALSE,
  fit = FALSE,
  delays = FALSE,
  overwrite = TRUE,
  posterior_nowcast = TRUE,
  posterior_R = TRUE,
  R_generative = TRUE, 
)
```

Fully stepwise approach
```{r}
model_def <- define_model(
  model_type = "impute_then_adjust",
  D = 7*8, # --> 8 weeks
  delay_changepoint = "segmented",
  n_lambda_pre = 0,
  overdispersion = TRUE,
  ts_model = "ets",
  sma_window = 0,
  ets_diff = FALSE,
  ets_noncentered = TRUE,
  profile = TRUE,
  threads = FALSE,
  force_recompile = FALSE)

prior_def <- define_priors(
  model_def,
  dirichlet_prior = FALSE,
  ets_alpha_fixed = 1,
  ets_beta_fixed = 1,
  ets_phi_fixed = 0,
  latent_delay_dist = latent_delay_dist_list,
  generation_time_dist = generation_time_dist_list,
  reporting_proportion = 0.02,
  alpha_logit_start = get_prior("alpha_logit_start", mu = 0, sd = 2)
  )

nowcast_job <- define_job(job_name = "real_fully_stepwise", data_def = data_def, model_def = model_def, prior_def = prior_def, sampling_def = sampling_def, output_def = output_def, index_by_date = T)
saveRDS(nowcast_job, here::here("results", "real_fully_stepwise", "real_fully_stepwise.rds"))
```

Fully generative approach
```{r}
model_def <- define_model(
  model_type = "impute_adjust_renewal",
  D = 7*8, # --> 8 weeks
  delay_changepoint = "segmented",
  n_lambda_pre = 0,
  overdispersion = TRUE,
  ts_model = "ets",
  sma_window = 0,
  ets_diff = FALSE,
  ets_noncentered = TRUE,
  profile = TRUE,
  threads = FALSE,
  force_recompile = FALSE)

prior_def <- define_priors(
  model_def,
  dirichlet_prior = FALSE,
  ets_alpha_fixed = 1,
  ets_beta_fixed = 1,
  ets_phi_fixed = 0,
  latent_delay_dist = latent_delay_dist_list,
  generation_time_dist = generation_time_dist_list,
  reporting_proportion = 0.02,
  alpha_logit_start = get_prior("alpha_logit_start", mu = 0, sd = 2),
  R_level_start = get_prior("R_level_start", mu = 2, sd = 0.8)
  )

nowcast_job <- define_job(job_name = "real_fully_generative", data_def = data_def, model_def = model_def, prior_def = prior_def, sampling_def = sampling_def, output_def = output_def, index_by_date = T)
saveRDS(nowcast_job, here::here("results", "real_fully_generative", "real_fully_generative.rds"))
```

Direct Rt estimation approach
```{r}
model_def <- define_model(
  model_type = "renewal_direct",
  D = 7*8, # --> 8 weeks
  delay_changepoint = "segmented",
  n_lambda_pre = 0,
  overdispersion = TRUE,
  ts_model = "ets",
  sma_window = 0,
  ets_diff = FALSE,
  ets_noncentered = TRUE,
  profile = TRUE,
  threads = FALSE,
  force_recompile = FALSE)

prior_def <- define_priors(
  model_def,
  dirichlet_prior = FALSE,
  ets_alpha_fixed = 1,
  ets_beta_fixed = 1,
  ets_phi_fixed = 0,
  latent_delay_dist = latent_delay_dist_list,
  generation_time_dist = generation_time_dist_list,
  reporting_proportion = 0.02,
  alpha_logit_start = get_prior("alpha_logit_start", mu = 0, sd = 2),
  R_level_start = get_prior("R_level_start", mu = 2, sd = 0.8)
  )

nowcast_job <- define_job(job_name = "real_direct", data_def = data_def, model_def = model_def, prior_def = prior_def, sampling_def = sampling_def, output_def = output_def, index_by_date = T)
saveRDS(nowcast_job, here::here("results", "real_direct", "real_direct.rds"))
```

## Running jobs

The nowcasting job arrays are stored in the "results" folder. Running all the jobs likely requires an HPC cluster. 
To run a job locally, first load the job array:
```{r}
jobname <- "real_direct"
jobarray <- readRDS(here::here("results", jobname, paste0(jobname, ".rds")))
```

Then run one of the jobs from the array:
```{r}
jobid <- 1
nowcast_result <- make_nowcast_from_job(jobarray, jobid)
```

Then save the result in a corresponding result file:
```{r}
output_filepath <- here::here(
  "results", jobname, paste0(jobname, "_", jobid, "_result", ".rds"))
print(paste("Saving results to", output_filepath))
saveRDS(nowcast_result, output_filepath)
```
