---
title: "Evaluation of nowcasting results on synthetic data"
---
# Preparation
```{r, message=F, output=F}
here::i_am("code/validation_synthetic.Rmd")
source(here::here("code", "utils", "setup.R"))
source(here::here("code", "utils", "utils_simulate.R"))
```

## Select wave
Please run either of the two chunks below to evaluate results for the first or second wave, respectively.

1st wave
```{r}
list_of_sources <- list(
  R_direct = "wave1_R_direct",
  R_stepwise = "wave1_R_stepwise",
  R_generative = "wave1_R_generative",
  miss_stepwise = "wave1_miss_stepwise",
  miss_stepwise_independent = "wave1_miss_stepwise_independent",
  miss_generative = "wave1_miss_generative",
  R_stepwise_ets = "wave1_R_stepwise_ets",
  R_generative_ets = "wave1_R_generative_ets"
)
ground_truth_sim <- readRDS(here::here("data", "simulated", "sim_wave1.rds"))

reference_date <- zoo::as.Date(0)

all_nowcast_dates <- reference_date + sort(unique(rep(c(70, 83, 104, 135),3) + rep(c(0,7,14), each = 4)))

feather_days_1st_wave <- c(`Before peak I` = 70, `At peak I` = 83, `After peak I` = 97, `Suppression` = 135)

phases_wave <- data.frame(
  occurrence_date = reference_date + feather_days_1st_wave,
  phase = forcats::fct_inorder(c("Before peak I", "At peak I", "After peak I", "Suppression"))
)
```

2nd wave
```{r}
list_of_sources <- list(
  R_direct = "wave2_R_direct",
  R_stepwise = "wave2_R_stepwise",
  R_generative = "wave2_R_generative",
  miss_stepwise = "wave2_miss_stepwise",
  miss_stepwise_independent = "wave2_miss_stepwise_independent",
  miss_generative = "wave2_miss_generative",
  R_stepwise_ets = "wave2_R_stepwise_ets",
  R_generative_ets = "wave2_R_generative_ets"
)
ground_truth_sim <- readRDS(here::here("data", "simulated", "sim_wave2.rds"))

reference_date <- zoo::as.Date(0)

all_nowcast_dates <- readRDS(here::here("results",list_of_sources[[1]],paste0(list_of_sources[[1]],".rds")))$data_def$now

feather_days_2nd_wave <- c(`Control` = 70, `Before peak II` =100, `At peak II` = 124, `After peak II` = 148)

phases_wave <- data.frame(
  occurrence_date = reference_date + feather_days_2nd_wave,
  phase = forcats::fct_inorder(c("Control","Before peak II","At peak II", "After peak II"))
)
```

## Get results
```{r}
res_list <- define_result_list(result_folder = "results", result_info = T, list_of_sources = list_of_sources, overwrite = F)
maxDelay <- max(sapply(res_list,function(x) max(x$maxDelay,na.rm=T)))
```

## Load nowcasting results
```{r}
n_res_list <- load_results_sim(res_list, maxDelay, reference_date, ground_truth_sim, keep_posterior = FALSE, overwrite = F, mc_cores = 6)
```

## Get diagnostics
```{r}
diags_list <- lapply(res_list, get_diagnostics, overwrite = F)
```

Check for model runs that need to be inspected due to diagnostics
```{r}
to_inspect <- lapply(diags_list, function(df) df %>% filter((chains_divergent > 0) | (chains_low_ebfmi > 0) | (ess_nowcast!="") | (ess_R!="") | (rhat_nowcast!="") | (rhat_R!="")))
to_inspect <- bind_rows(to_inspect, .id = "approach")
to_inspect
```

Check rhat warnings for R: inspect respective model runs for potentially biased R estimation (none found)
```{r}
apply(to_inspect, 1, function(x) {
  inspect <- n_res_list[[x["approach"]]] |>
    filter(id %in% as.integer(unique(x["id"]))) |>
    mutate(approach = x["approach"]) |> 
    select(approach, R_model, id, nowcast_date, dataset_index, date_index, delay, date, starts_with("R")) |>
    arrange(id, dataset_index, date_index, delay)
  if (nrow(inspect) > 0) {
    return(inspect)
  } else {
    return(data.frame())
  }
})
```

## Compute performance metrics
```{r}
nowcast_timing_data <- data.frame(
  difference = c(0, 7, 7, 14, 21),
  nowcast_timing = forcats::fct_inorder(c("0-6", "7-13", "8-13", "14-20", "21-27"))
  ) %>%
  mutate(nowcast_timing_list = forcats::fct_inorder(range_to_comma_list(nowcast_timing), ordered = T)) %>% 
  full_join(data.frame(delay = 0:6), by = character())

nowcast_timing_data[nowcast_timing_data$nowcast_timing=="8-13","delay"] <- nowcast_timing_data[nowcast_timing_data$nowcast_timing=="8-13","delay"] + 1
nowcast_timing_data <- nowcast_timing_data %>% filter(delay<6)

phases <- phases_wave %>%
  full_join(nowcast_timing_data, by = character()) %>%
  mutate(
    nowcast_date = occurrence_date + difference,
    delay = as.integer(delay + difference)
  ) %>%
  select(phase, occurrence_date, nowcast_timing, nowcast_timing_list, nowcast_date, delay)

table_timings <- c(
    "In same week" =  "0-6",
    "One week after" = "7-13",
    "Two weeks after" = "14-20"
  )
```

```{r}
m_res_list <- lapply(n_res_list, function(res) get_metrics(res %>% mutate(delay = as.integer(delay)) %>% right_join(phases, by = c("nowcast_date", "delay")) %>% group_by(dataset_index, R_model, phase, nowcast_timing, nowcast_timing_list)))
```

# Results

## Plot preparation
```{r}
model_metainf <- tribble(
  ~ internal_name, ~label, ~label_short, ~label_short_1st, ~label_short_2nd, ~ label_long, ~ color,
  "R_direct renewal",
  "Direct (renewal model)",
  "Direct", "Direct", "",
  "Complete data:\nDirect (renewal model)", "#ecb613",
  
  "R_stepwise renewal",
  "Stepwise (truncation adjustment -> renewal model)",
  "Stepwise", "Stepwise", "",
  "Complete data:\nStepwise (renewal model)", "#4DBBD5FF",
  
  "R_stepwise epiestim",
  "Stepwise (truncation adjustment -> EpiEstim)",
  "Stepwise (EpiEstim)", "Stepwise", "(EpiEstim)",
  "Complete data:\nStepwise (EpiEstim)", "#009933",
  
  "R_generative renewal",
  "Generative (truncation adjustment + renewal model)",
  "Generative", "Generative", "",
  "Complete data:\nGenerative (renewal model)", "#E64B35FF",
  
  "miss_stepwise_independent renewal",
  "Stepwise (independent imputation -> nowcasting model)",
  "Stepwise (independent)", "Stepwise", "(independent)", 
  "Incomplete data:\nStepwise (independent imputation)", "#85e085",
  
  "miss_stepwise renewal",
  "Stepwise (backward imputation -> nowcasting model)",
  "Stepwise (backward)", "Stepwise", "(backward)",
  "Incomplete data:\nStepwise (backward imputation)", "#3C5488FF",
  
  "miss_generative renewal",
  "Generative (missingness + nowcasting model)",
  "Generative",  "Generative", "",
  "Incomplete data:\nGenerative (missingness model)", "#F39B7FFF",
  
  "R_stepwise_ets renewal",
  "Stepwise (truncation adjustment (ETS) -> renewal model)",
  "Stepwise (non-stationary)", "Stepwise", "(non-stationary)",
  "Complete data:\nStepwise (renewal model)", "#4d91d5",
  
  "R_stepwise_ets epiestim",
  "Stepwise (truncation adjustment (ETS) -> EpiEstim)",
  "Stepwise (non-stationary, EpiEstim)", "Stepwise", "(non-stationary, EpiEstim)",
  "Complete data:\nStepwise (EpiEstim)", "#00A087FF",
  
  "R_generative_ets renewal",
  "Generative (truncation adjustment + renewal model (ETS))",
  "Generative (non-stationary)", "Generative", "(non-stationary)",
  "Complete data:\nGenerative (renewal model)", "#9b9b4b",
)

model_names <- setNames(model_metainf$label, model_metainf$internal_name)
model_names_short <- setNames(model_metainf$label_short, model_metainf$internal_name)
model_names_short_1st <- setNames(model_metainf$label_short_1st, model_metainf$internal_name)
model_names_short_2nd <- setNames(model_metainf$label_short_2nd, model_metainf$internal_name)
model_colors <- setNames(model_metainf$color, model_metainf$label)
```

```{r}
source(here::here("code","utils","plot_legends_synthetic.R"))
```

## First wave

### Main
```{r, fig.width = 10, fig.height = 12}
select_dataset <- 7
save_plots <- T
make_plots <- T
save_tables <- T

suppressWarnings({

  # Nowcast cases, R models
  if (make_plots) plot_v <- plot_validation_select(dataset_id = select_dataset,
                      models = c("R_stepwise", "R_generative"),
                      R_models = "renewal",
                      feather_days = feather_days_1st_wave,
                      feather_maxdelay = 7*2+3,
                      feather_margin = 18,
                      plot_type = "Cases",
                      percent_best = T,
                      categories_name = "Nowcasting approach")
  if (save_plots) {
    ggsave(here::here("figures","paper","wave1_feather_renewal_Nt.pdf"), plot = plot_v, width = 10, height = 12)
    ragg_save_TIFF(the_plot = plot_v, here::here("figures","paper","wave1_feather_renewal_Nt.tiff"), width = 5, height = 6, units = "in", pointsize = 20, res = 500, scaling = 0.5, compression = "lzw", background = "white")
  }
  if (save_tables) {
      make_performance_table_latex(
        outcome_name = "nowcast_all",
        models = c("R_stepwise", "R_generative"),
        R_models = "renewal",
        timings = table_timings
        ) |> 
      write_file(
        file = here::here("tables","paper","wave1_feather_renewal_Nt.tex")
        )
  }

  # Nowcast Rt, R models
  if (make_plots) plot_v <- plot_validation_select(dataset_id = select_dataset,
                      models = c("R_direct", "R_stepwise", "R_generative"),
                      R_models = "renewal",
                      feather_days = feather_days_1st_wave,
                      feather_maxdelay = 7*2+3,
                      feather_margin = 18,
                      plot_type = "R",
                      percent_best = T,
                      categories_name = "Nowcasting approach")
  if (save_plots) {
    ggsave(here::here("figures","paper","wave1_feather_renewal_Rt.pdf"), plot = plot_v, width = 10, height = 12)
    ragg_save_TIFF(the_plot = plot_v, here::here("figures","paper","wave1_feather_renewal_Rt.tiff"), width = 5, height = 6, units = "in", pointsize = 20, res = 500, scaling = 0.5, compression = "lzw", background = "white")
  }
  if (save_tables) {
    make_performance_table_latex(
      outcome_name = "R",
      models = c("R_direct", "R_stepwise", "R_generative"),
      R_models = "renewal",
      timings = table_timings
      ) |> 
    write_file(
      file = here::here("tables","paper","wave1_feather_renewal_Rt.tex")
    )
  }

  # Nowcast cases, missingness models
  if (make_plots) plot_v <- plot_validation_select(dataset_id = select_dataset,
                      models = c("miss_stepwise_independent", "miss_stepwise", "miss_generative"),
                      feather_days = feather_days_1st_wave,
                      feather_maxdelay = 7*2+3,
                      feather_margin = 18,
                      plot_type = "Cases",
                      percent_best = T,
                      legends.rel_widths = c(0.3, 0.5, 0.2),
                      categories_name = "Imputation approach")
  if (save_plots) {
    ggsave(here::here("figures","paper","wave1_feather_miss_Nt.pdf"), plot = plot_v, width = 10, height = 12)
    ragg_save_TIFF(the_plot = plot_v, here::here("figures","paper","wave1_feather_miss_Nt.tiff"), width = 5, height = 6, units = "in", pointsize = 20, res = 500, scaling = 0.5, compression = "lzw", background = "white")
  }
  if (save_tables) {
      make_performance_table_latex(
        outcome_name = "nowcast_all",
        models = c("miss_stepwise_independent", "miss_stepwise", "miss_generative"),
        R_models = "renewal",
        timings = table_timings
        ) |> 
      write_file(
        file = here::here("tables","paper","wave1_feather_miss_Nt.tex")
        )
  }

  # Nowcast Rt, missingness models
  if (make_plots) plot_v <- plot_validation_select(dataset_id = select_dataset,
                      models = c("miss_stepwise_independent", "miss_stepwise", "miss_generative"),
                      feather_days = feather_days_1st_wave,
                      feather_maxdelay = 7*2+3,
                      feather_margin = 18,
                      plot_type = "R",
                      percent_best = T,
                      categories_name = "Imputation approach")
  if (save_plots) {
    ggsave(here::here("figures","paper","wave1_feather_miss_Rt.pdf"), plot = plot_v, width = 10, height = 12)
    ragg_save_TIFF(the_plot = plot_v, here::here("figures","paper","wave1_feather_miss_Rt.tiff"), width = 5, height = 6, units = "in", pointsize = 20, res = 500, scaling = 0.5, compression = "lzw", background = "white")
  }
  if (save_tables) {
      make_performance_table_latex(
        outcome_name = "R",
        models = c("miss_stepwise_independent", "miss_stepwise", "miss_generative"),
        R_models = "renewal",
        timings = table_timings
        ) |> 
      write_file(
        file = here::here("tables","paper","wave1_feather_miss_Rt.tex")
        )
  }
})
```

### ETS models
```{r, fig.width = 10, fig.height = 12}
select_dataset <- 7
save_plots <- T
make_plots <- T
save_tables <- T

# Nowcast cases
suppressWarnings({
if (make_plots) plot_validation_select(dataset_id = select_dataset,
                    models = c("R_generative", "R_stepwise_ets", "R_generative_ets"),
                    R_models = "renewal",
                    feather_days = feather_days_1st_wave,
                    feather_maxdelay = 7*2+3,
                    feather_margin = 18,
                    plot_type = "Cases",
                    percent_best = T,
                    categories_name = "Nowcasting approach")})
if (save_plots) ggsave(here::here("figures","paper","wave1_feather_renewal_ets_Nt.pdf"), width = 10, height = 12)
if (save_tables) {
      make_performance_table_latex(
        outcome_name = "nowcast_all",
        models = c("R_generative", "R_stepwise_ets", "R_generative_ets"),
        R_models = "renewal",
        timings = table_timings
        ) |> 
      write_file(
        file = here::here("tables","paper","wave1_feather_renewal_ets_Nt.tex")
        )
}

# Nowcast Rt
suppressWarnings({
if (make_plots) plot_validation_select(dataset_id = select_dataset,
                    models = c("R_generative", "R_stepwise_ets", "R_generative_ets"),
                    R_models = "renewal",
                    feather_days = feather_days_1st_wave,
                    feather_maxdelay = 7*2+3,
                    feather_margin = 18,
                    plot_type = "R",
                    percent_best = T,
                    categories_name = "Nowcasting approach")})
if (save_plots) ggsave(here::here("figures","paper","wave1_feather_renewal_ets_Rt.pdf"), width = 10, height = 12)
if (save_tables) {
      make_performance_table_latex(
        outcome_name = "R",
        models = c("R_generative", "R_stepwise_ets", "R_generative_ets"),
        R_models = "renewal",
        timings = table_timings
        ) |> 
      write_file(
        file = here::here("tables","paper","wave1_feather_renewal_ets_Rt.tex")
        )
}

```

### EpiEstim
```{r, fig.width = 10, fig.height = 8.5}
select_dataset <- 7
save_plots <- T
make_plots <- T
save_tables <- T

if (make_plots) plot_validation_select(
  dataset_id = select_dataset,
  models = c("R_stepwise"),
  R_models = c("renewal", "epiestim"),
  feather_days = feather_days_1st_wave,
  feather_maxdelay = 7 * 2 + 3,
  lags_list = list(
    "Nowcasts one week after" = 8:13,
    "Nowcasts two weeks after" = 14:20
  ),
  feather_margin = 18,
  plot_type = "R",
  percent_best = T,
  legend_plot.rel_heights = c(1, 6.3),
  categories_name = "Nowcasting approach"
)
if (save_plots) ggsave(here::here("figures", "paper", "wave1_feather_epiestim_Rt.pdf"), width = 10, height = 8.5)
if (save_tables) {
      make_performance_table_latex(
        outcome_name = "R",
        models = c("R_stepwise"),
        R_models = c("renewal", "epiestim"),
        timings = c(
          "One week after" = "8-13",
          "Two weeks after" = "14-20"
          )
        ) |> 
      write_file(
        file = here::here("tables","paper","wave1_feather_epiestim_Rt.tex")
        )
}
```

```{r}
make_performance_table(
        outcome_name = "R",
        models = c("R_stepwise"),
        R_models = c("renewal", "epiestim"),
        timings = c(
          "One week after" = "8-13",
          "Two weeks after" = "14-20"
          )
        ) |>
  mutate(wis_rel = `R_mean_wis_Stepwise (EpiEstim)`/R_mean_wis_Stepwise) |> 
  pull(wis_rel) |> dput()
```


## Second wave

### Main
```{r, fig.width = 10, fig.height = 12}
select_dataset <- 2
save_plots <- T
make_plots <- T
save_tables <- T

suppressWarnings({
  # Nowcast cases, R models
  if (make_plots) plot_validation_select(dataset_id = select_dataset,
                      models = c("R_stepwise", "R_generative"),
                      R_models = "renewal",
                      feather_days = feather_days_2nd_wave,
                      feather_maxdelay = 7*2+3,
                      feather_margin = 18,
                      plot_type = "Cases",
                      percent_best  = T,
                      categories_name = "Nowcasting approach")
  if (save_plots) ggsave(here::here("figures","paper","wave2_feather_renewal_Nt.pdf"), width = 10, height = 12)
    if (save_tables) {
      make_performance_table_latex(
        outcome_name = "nowcast_all",
        models = c("R_stepwise", "R_generative"),
        R_models = "renewal",
        timings = table_timings
        ) |> 
      write_file(
        file = here::here("tables","paper","wave2_feather_renewal_Nt.tex")
        )
  }

  # Nowcast Rt, R models
  if (make_plots) plot_validation_select(dataset_id = select_dataset,
                      models = c("R_direct", "R_stepwise", "R_generative"),
                      R_models = "renewal",
                      feather_days = feather_days_2nd_wave,
                      feather_maxdelay = 7*2+3,
                      feather_margin = 18,
                      plot_type = "R",
                      percent_best = T,
                      categories_name = "Nowcasting approach")
  if (save_plots) ggsave(here::here("figures","paper","wave2_feather_renewal_Rt.pdf"), width = 10, height = 12)
  if (save_tables) {
    make_performance_table_latex(
      outcome_name = "R",
      models = c("R_direct", "R_stepwise", "R_generative"),
      R_models = "renewal",
      timings = table_timings
      ) |> 
    write_file(
      file = here::here("tables","paper","wave2_feather_renewal_Rt.tex")
    )
  }

  # Nowcast cases, missingness models
  if (make_plots) plot_validation_select(dataset_id = select_dataset,
                      models = c("miss_stepwise_independent", "miss_stepwise", "miss_generative"),
                      feather_days = feather_days_2nd_wave,
                      feather_maxdelay = 7*2+3,
                      feather_margin = 18,
                      plot_type = "Cases",
                      percent_best = T,
                      legends.rel_widths = c(0.3, 0.5, 0.2),
                      categories_name = "Imputation approach")
  if (save_plots) ggsave(here::here("figures","paper","wave2_feather_miss_Nt.pdf"), width = 10, height = 12)
  if (save_tables) {
      make_performance_table_latex(
        outcome_name = "nowcast_all",
        models = c("miss_stepwise_independent", "miss_stepwise", "miss_generative"),
        R_models = "renewal",
        timings = table_timings
        ) |> 
      write_file(
        file = here::here("tables","paper","wave2_feather_miss_Nt.tex")
        )
  }

  # Nowcast Rt, missingness models
  if (make_plots) plot_validation_select(dataset_id = select_dataset,
                      models = c("miss_stepwise_independent", "miss_stepwise", "miss_generative"),
                      feather_days = feather_days_2nd_wave,
                      feather_maxdelay = 7*2+3,
                      feather_margin = 18,
                      plot_type = "R",
                      percent_best = T,
                      categories_name = "Imputation approach")
  if (save_plots) ggsave(here::here("figures","paper","wave2_feather_miss_Rt.pdf"), width = 10, height = 12)
  if (save_tables) {
      make_performance_table_latex(
        outcome_name = "R",
        models = c("miss_stepwise_independent", "miss_stepwise", "miss_generative"),
        R_models = "renewal",
        timings = table_timings
        ) |> 
      write_file(
        file = here::here("tables","paper","wave2_feather_miss_Rt.tex")
        )
  }
  
})
```

### ETS models
```{r, fig.width = 10, fig.height = 12}
select_dataset <- 2
save_plots <- T
make_plots <- T
save_tables <- T

# Nowcast cases
if (make_plots) plot_validation_select(dataset_id = select_dataset,
                    models = c("R_generative", "R_stepwise_ets", "R_generative_ets"),
                    R_models = "renewal",
                    feather_days = feather_days_2nd_wave,
                    feather_maxdelay = 7*2+3,
                    feather_margin = 18,
                    plot_type = "Cases",
                    percent_best = T, 
                    categories_name = "Nowcasting approach")
if (save_plots) ggsave(here::here("figures","paper","wave2_feather_renewal_ets_Nt.pdf"), width = 10, height = 12)
if (save_tables) {
      make_performance_table_latex(
        outcome_name = "nowcast_all",
        models = c("R_generative", "R_stepwise_ets", "R_generative_ets"),
        R_models = "renewal",
        timings = table_timings
        ) |> 
      write_file(
        file = here::here("tables","paper","wave2_feather_renewal_ets_Nt.tex")
        )
}

# Nowcast Rt
if (make_plots) plot_validation_select(dataset_id = select_dataset,
                    models = c("R_stepwise_ets", "R_generative_ets", "R_generative"),
                    R_models = "renewal",
                    feather_days = feather_days_2nd_wave,
                    feather_maxdelay = 7*2+3,
                    feather_margin = 18,
                    plot_type = "R",
                    percent_best = T,
                    categories_name = "Nowcasting approach")
if (save_plots) ggsave(here::here("figures","paper","wave2_feather_renewal_ets_Rt.pdf"), width = 10, height = 12)
if (save_tables) {
      make_performance_table_latex(
        outcome_name = "R",
        models = c("R_generative", "R_stepwise_ets", "R_generative_ets"),
        R_models = "renewal",
        timings = table_timings
        ) |> 
      write_file(
        file = here::here("tables","paper","wave2_feather_renewal_ets_Rt.tex")
        )
}

```

### EpiEstim
```{r, fig.width = 10, fig.height = 8.5}
select_dataset <- 2
save_plots <- T
make_plots <- T
save_tables <- T

if (make_plots) plot_validation_select(
  dataset_id = select_dataset,
  models = c("R_stepwise"),
  R_models = c("renewal", "epiestim"),
  feather_days = feather_days_2nd_wave,
  feather_maxdelay = 7 * 2 + 3,
  lags_list = list(
    "Nowcasts one week after" = 8:13,
    "Nowcasts two weeks after" = 14:20
  ),
  feather_margin = 18,
  plot_type = "R",
  percent_best = T,
  legend_plot.rel_heights = c(1, 6.3),
  categories_name = "Nowcasting approach"
)
if (save_plots) ggsave(here::here("figures", "paper", "wave2_feather_epiestim_Rt.pdf"), width = 10, height = 8.5)
if (save_tables) {
      make_performance_table_latex(
        outcome_name = "R",
        models = c("R_stepwise"),
        R_models = c("renewal", "epiestim"),
        timings = c(
          "One week after" = "8-13",
          "Two weeks after" = "14-20"
          )
        ) |> 
      write_file(
        file = here::here("tables","paper","wave2_feather_epiestim_Rt.tex")
        )
}
```
```{r}
make_performance_table(
        outcome_name = "R",
        models = c("R_stepwise"),
        R_models = c("renewal", "epiestim"),
        timings = c(
          "One week after" = "8-13",
          "Two weeks after" = "14-20"
          )
        ) |> 
  mutate(wis_rel = `R_mean_wis_Stepwise (EpiEstim)`/R_mean_wis_Stepwise) |> 
  pull(wis_rel) |> dput()
```
