---
title: "Simulate 2nd wave"
---

```{r}
here::i_am("code/simulations/simulate_wave2.R")
source(here::here("code", "utils", "utils.R"))
source(here::here("code", "utils", "utils_simulate.R"))
library(extraDistr)
library(plotly)
library(tibble)
library(dplyr)
library(tidyr)
```

```{r}
n_realizations <- 50
```

## Generation time distribution
```{r}
maxGen <- 21
gen_wildtype <- get_discrete_gamma(gamma_mean = 4.9, gamma_sd = 4.1, maxX = maxGen, include_zero = F)

generation_dist <- rbind(
  # initial infections (length maxGen)
  get_distribution_time_series(
    tribble_format = tribble(
      ~changepoints, ~pmf_list,
      1, gen_wildtype,
      maxGen, gen_wildtype
    )
  ),
  # modeled infections
  get_distribution_time_series(
    tribble_format = tribble(
      ~changepoints, ~pmf_list,
      1, gen_wildtype,
      260, gen_wildtype
    )
  )
)
plot_dist_ts(generation_dist, ylab = "Generation time")
plot_scalar_ts(dist_get_mean(generation_dist))
```
## Incubation period distribution
```{r}
maxInc <- 21
inc_wildtype <- get_discrete_gamma(gamma_mean = 5.3, gamma_sd = 3.2, maxX = maxInc)

incubation_dist <- get_distribution_time_series(
  tribble_format = tribble(
    ~changepoints, ~pmf_list,
    1, inc_wildtype,
    260, inc_wildtype
  )
)
plot_dist_ts(incubation_dist, ylab = "Incubation time")
plot_scalar_ts(dist_get_mean(incubation_dist))
```
## Initial infections
```{r}
initial_infections <- replicate(n_realizations, rpois(maxGen, 1000), simplify = F)
plot_scalar_ts(ts = initial_infections[[1]], ylab = "Initial infections", xlab = "Pre-time")
```
## Reproduction number
```{r}
R <- get_scalar_time_series(tribble_format = tribble(
  ~changepoints, ~scalars,
  1, 1,
  70, 1,
  80, 1.4,
  100, 1.4,
  130, 0.7,
  170, 0.7,
))
plot_scalar_ts(R, ylab = "Effective reproduction number")
```
## Infection trajectory
```{r}
set.seed(0)
infections <- list() 
for (i in 1:n_realizations) {
  print(paste("Simulate scenario",i))
  inf_sim <- simulate_infections(initial_infections[[i]], R, generation_dist)
  infections[[i]] <- inf_sim
  print(paste("Maximum number of infections:",max_inf))
}
(qplot(y=infections[[3]][["I"]]) + geom_line(y=infections[[3]][["I_expected"]]) + ylab("Infections") + theme_bw()) #%>% ggplotly()
```
## Probability of symptom onsets to be known (converse of missingness probability)
```{r}
# roughly between 40% and 80% (i.e. 20%-60% missingness)
onset_known_prob <- replicate(n_realizations, get_scalar_time_series(changepoints = seq(1,300,by=28), scalars= 0.4 + 0.4*plogis(cumsum(rnorm(length(seq(1,300,by=28)),0,0.5)))), simplify = F)

plot_scalar_ts(onset_known_prob[[3]], ylab = "Share of known symptom onsets", xlab = "Time") + coord_cartesian(ylim=c(0.35,0.85))
```
## Hospitalization delay distribution (symptom onset to hospitalization)
Only relevant if nowcasting by date of hospitalization instead of symptom onset
```{r}
maxSympToHosp <- 25
symp_to_hosp_dist <- get_distribution_time_series(
  tribble_format = tribble(
    ~changepoints, ~pmf_list,
    1, get_discrete_lognormal(0.8, 0.3, maxX = maxSympToHosp),
    60, get_discrete_lognormal(0.9, 0.325, maxX = maxSympToHosp),
    200, get_discrete_lognormal(0.8, 0.35, maxX = maxSympToHosp),
    220, get_discrete_lognormal(0.9, 0.3, maxX = maxSympToHosp),
    355, get_discrete_lognormal(0.95, 0.4, maxX = maxSympToHosp),
  )
)
plot_scalar_ts(dist_get_mean(symp_to_hosp_dist), ylab = "Mean delay")
plot_dist_ts(symp_to_hosp_dist, ylab = "Time from symptom onset to hospitalization")
plot(symp_to_hosp_dist[10, ])
```
## Reporting delay distribution (symptom onset to report)
```{r}
base_p <- get_discrete_lognormal(unit_mean = 9, unit_sd = 8, maxX = 7*14) # in line with real Swiss data

data.frame(p = base_p) %>%
  mutate(delay = 0:(n() - 1)) %>%
  ggplot(aes(x = delay, y = p)) +
  geom_bar(stat = "identity") +
  theme_bw()
data.frame(p = base_p) %>%
  mutate(delay = 0:(n() - 1)) %>%
  ggplot(aes(x = delay, y = cumsum(p))) +
  geom_bar(stat = "identity") +
  theme_bw()

symp_to_rep_dist <- replicate(n_realizations, get_p_time_series(
  qlogis(get_hazard_from_p(base_p)),
  changepoints = seq(1,300,by=28),
  scalars = -0.3 + 0.6*plogis(cumsum(rnorm(length(seq(1,300,by=28)),0,0.5))), #c(0, -0.1, -0.2, -0.2, -0.1, 0.1, 0.15, 0.2, 0.35, 0.4, 0.3),
  rep_seasonality = c(0.1, 0.05, 0, 0, 0, -0.7, -0.8)
), simplify = F)

plot_scalar_ts(dist_get_mean(symp_to_rep_dist[[1]]), ylab = "Mean delay")

data.frame(p = symp_to_rep_dist[[1]][1, ]) %>%
  mutate(delay = 0:(n() - 1)) %>%
  ggplot(aes(x = delay, y = p)) +
  geom_bar(stat = "identity") +
  theme_bw()
data.frame(p = symp_to_rep_dist[[1]][100, ]) %>%
  mutate(delay = 0:(n() - 1)) %>%
  ggplot(aes(x = delay, y = p)) +
  geom_bar(stat = "identity") +
  theme_bw()
data.frame(p = symp_to_rep_dist[[1]][153, ]) %>%
  mutate(delay = 0:(n() - 1)) %>%
  ggplot(aes(x = delay, y = p)) +
  geom_bar(stat = "identity") +
  theme_bw()
plot_dist_ts(symp_to_rep_dist[[1]], ylab = "Time from symptom onset to reporting")
```
## Ascertainment / reporting proportion
```{r}
hosp_to_rep_dist <- NULL
hosp_prob = c(rep(0.02, 200))
```

## Simulate
```{r}
cases <- lapply(1:n_realizations,function(i) simulate_outcomes_cases(infections[[i]][["I"]], incubation_dist, hosp_prob, symp_to_hosp_dist))
cases <- lapply(1:n_realizations,function(i) simulate_ascertainment_cases(cases[[i]], onset_known_prob[[i]], symp_to_rep_dist[[i]], hosp_to_rep_dist))
```

```{r}
sim_def <- lapply(1:n_realizations, function(i) define_simulation(
  initial_infections = initial_infections[[i]],
  R = R,
  generation_dist = generation_dist,
  incubation_dist = incubation_dist,
  hosp_prob = hosp_prob,
  symp_to_hosp_dist = symp_to_hosp_dist,
  onset_known_prob = onset_known_prob[[i]],
  symp_to_rep_dist = symp_to_rep_dist[[i]],
  hosp_to_rep_dist = hosp_to_rep_dist
))

sim <- list()
for (i in 1:n_realizations) {
  process_summary <- summarize_linelist(cases[[i]], 1:length(infections[[i]][["I"]]))
  process_summary[["infections"]] <- infections[[i]][["I"]]
  process_summary[["infections_expected"]] <- infections[[i]][["I_expected"]]
  process_summary[["t_onset"]] <- 1:(length(infections[[i]][["I"]]) + ncol(incubation_dist) - 1)
  sim[[i]] <- list(parameters = sim_def[[i]], linelist = cases[[i]], process_summary = process_summary)
}

```

```{r}
saveRDS(sim, here::here("data", "simulated", "sim_wave2.rds"))
```

```{r}
sim <- readRDS(here::here("data", "simulated", "sim_wave2.rds"))
```

## Inspect simulation results
```{r}
sim_index <- 1
(sim[[sim_index]][["linelist"]] %>%
  ggplot() +
  geom_line(data = sim[["linelist"]], aes(x = infection_time), stat = "count", color = "orange") +
  geom_line(data = sim[["linelist"]], aes(x = onset_time), stat = "count", color = "red") +
  geom_line(data = sim[[sim_index]][["linelist"]] %>% filter(onset_known), aes(x = onset_time), stat = "count", color = "red", linetype = "dashed") +
  geom_line(aes(x = hosp_time), stat = "count", color = "blue") +
  geom_line(data = sim[[sim_index]][["linelist"]] %>% filter(onset_known), aes(x = hosp_time), stat = "count", color = "blue", linetype = "dashed") +
  geom_line(aes(x = rep_time), stat = "count", color = "black") +
  geom_line(data = sim[[sim_index]][["linelist"]] %>% filter(onset_known), aes(x = rep_time), stat = "count", color = "black", linetype = "dashed") +
  theme_bw() +
  xlab("Time") +
  ylab("Count") +
  scale_x_continuous(expand = c(0, 0)) +
  coord_cartesian(xlim = c(0, max(sim[[sim_index]][["linelist"]]$infection_time))))
```

```{r, fig.width = 10, fig.height = 5}
sim_index <- 3
sim[[sim_index]][["linelist"]] %>%
    pivot_longer(cols = c(infection_time, onset_time, hosp_time, rep_time), names_to = "Event", values_to = "time") %>% 
    count(Event, time) %>% mutate(Event = recode_factor(Event,
                                               infection_time = "Infection",
                                               onset_time = "Symptom onset",
                                               hosp_time = "Hospitalisation",
                                               rep_time = "Report", .ordered = T)) %>% 
  filter(Event != "Hospitalisation") %>% 
  group_by(Event) %>% 
  complete(time = 0:200, fill = list(n = 0)) %>% 
  mutate(time = zoo::as.Date(0) + time) %>% 
  ggplot() +
  geom_line(aes(x = time, y = n, color = Event)) +
  theme_bw() +
  theme(legend.position = "none", strip.background = element_blank(), strip.text = element_text(hjust = 0)) +
  xlab("Date") +
  ylab("Number of cases") +
  scale_x_date(expand = c(0, 0)) +
  scale_color_manual(values = c("red", "blue", "black")) +
  coord_cartesian(xlim = zoo::as.Date(0) + c(30, 200)) +
  facet_wrap(~Event, ncol = 1)

ggsave(here::here("figures", "paper", "sim_wave2_example_trajectory.pdf"), width = 10, height = 5)
```

```{r}
sim_index <- 3
# control, before peak, at peak, after peak
nowcasting_dates <- c(70, 100, 124, 148)

plot_scalar_ts(sim[[sim_index]]$parameters$R, ylab = "Effective reproduction number") +
  geom_vline(data=data.frame(x=nowcasting_dates), aes(xintercept=x), linetype="dashed") +
geom_vline(data=data.frame(x=nowcasting_dates-7), aes(xintercept=x), linetype="dashed", linewidth = 0.2, color = "blue")

sim[[sim_index]][["linelist"]] %>%
  ggplot() +
  geom_line(aes(x = onset_time), stat = "count", color = "blue") +
  theme_bw() +
  xlab("Time") +
  ylab("Count") +
  scale_x_continuous(expand = c(0, 0)) +
  coord_cartesian(xlim = c(0, max(sim[[sim_index]][["linelist"]]$infection_time)), ylim = c(0, NA)) +
   geom_vline(data=data.frame(x=nowcasting_dates), aes(xintercept=x), linetype="dashed")
```

